---
title: "Observed COVID-19 Prevalence in the Commonwealth of Virginia"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Observed COVID-19 Prevalence in the Commonwealth of Virginia}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(vacovdata)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(latex2exp)
```

## Raw prevalence observations

These are the raw positive test fractions, statewide, uncorrected for enrichment 
due to targeted testing (what we call $\hat{f}$ in our paper).   

```{r rawobs}
pltdata <- filter(vacovdata, date > as.Date('2020-03-15'), !is.na(fpos)) %>%
  select(date, fpos, npos=positiveIncrease, nneg=negativeIncrease) %>%
  mutate(flo = qbeta(0.025, npos+1, nneg+1), fhi = qbeta(0.975, npos+1, nneg+1))

ggplot(pltdata, aes(x=date)) +
  geom_point(aes(y=fpos), size=1.5) +
  geom_errorbar(aes(ymin=flo, ymax=fhi)) +
  ylab(TeX('$\\hat{f}$')) +
  theme_bw()
```

Note, however, that there are several days in May that saw dramatic increases in
the number of tests performed.

```{r ntest}
pltdata <- mutate(pltdata, ntest = npos+nneg)
## Will use these May correlations below
maydata <- filter(pltdata, date >= as.Date('2020-05-01'))
maycor <- signif(cor(maydata$ntest, maydata$fpos, method='spearman'),2)

ggplot(pltdata, aes(x=date, y=ntest)) +
  geom_line(size=1.2) +
  ylab('Number of tests') +
  theme_bw()
```

These excess tests, relative to the surrounding days, tend to be negative, producing
low outlier points in the plot of $\hat{f}$, particularly on May 1st, 15th, 20th,
and 24th.  Overall for the month of May, the correlation between the number of
tests performed and the estimate of $\hat{f}$ (measured by the Spearman correlation
coefficient, $\rho_s$) is $\rho_s = `r maycor`$.  This high correlation suggests
that on the days with a surge in testing, a different selection procedure is being
used, resulting in a different enrichment factor.

## Growth rates

Both the positive test fraction and the growth rates derived from it are quite 
noisy, so we use a 5-day sliding window to accumulate observations.

```{r growthplt}
grpltdata <- select(vacovdata, date, pos=positive, tot=totalTestResults) %>%
  fitgr(7) %>% 
  filter(date > as.Date('2020-03-23'), !is.na(kobs))

ggplot(grpltdata, aes(x=date)) +
  geom_point(aes(y=kobs, color='uncorrected'), size=1.5) +
  geom_smooth(aes(y=kobs, color='uncorrected'), se=FALSE) +
  geom_point(aes(y=kpop, color='corrected'), size=1.5) +
  geom_smooth(aes(y=kpop, color='corrected'), se=FALSE) +
  #coord_cartesian(ylim=c(-0.1,0.3)) +
  #xlim(as.Date(c('2020-04-01', '2020-05-30'))) +
  scale_color_solarized() +
  theme_bw()
```

